---
title: Hindsight is 2020
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    orientation: rows
    theme: "journal"
runtime: shiny
css: custom.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# I loaded all of my packages and data in the helper.R file

source("subset_helper.R")

sims <- read_csv("app-data/simulations_subset.csv")

```

Home
================================================

### About

This interactive tool aims to supplement a semester-long project in which I forecasted the 2020 Presidential Election. The below interactive tools present and contextualize my predictions in a more visual manner than I did in my blog posts. Nate Silver, if you're seeing this, please hire me! See my full election blog and forecast write-up [here](https://kayla-manning.github.io/gov1347/). If you would like to take a closer look at the code to produce this app, feel free to poke around [this](https://github.com/kayla-manning/presidential-forecast-analysis) Github repository. The code for the original forecast is located at [this](https://github.com/kayla-manning/gov1347) repository.

Row
-------------------------------------------------

### Actual Electoral Votes

```{r actual_ev_gauge}

# Biden actual EV gauge

renderGauge(

  gauge(306,
        min = 0, 
        max = 538,
        gaugeSectors(
          success = c(270, 538), 
          warning = c(269, 269), 
          danger = c(0, 268)), 
        label = "Electoral Votes \nfor Biden")
)

# Trump actual EV gauge

renderGauge(
  
  gauge(232, 
        min = 0, 
        max = 538,
        gaugeSectors(
          success = c(270, 538),
          warning = c(269, 269), 
          danger = c(0, 268)),
        label = "Electoral Votes \nfor Trump")
)

```

### Actual Two-Party Popular Vote Share

```{r actual_pv2p_gauge}

# getting Biden's actual pv2p so that I can reference it when I make the gauges

biden_actual_pv2p <- enos_data %>% 
  summarise(democrat = sum(democrat),
            republican = sum(republican)) %>% 
  summarise(democrat = round(democrat / (democrat + republican) * 100, 3)) %>% 
  pull(democrat)

# Biden actual pv2p gauge

renderGauge(
  
  gauge(biden_actual_pv2p,
        min = 0, 
        max = 100,
        gaugeSectors(
          success = c(50.1, 100), 
          warning = c(50, 50), 
          danger = c(0, 49.9)), 
        label = "Two-Party Popular Vote \nfor Biden")
  
)

# Trump actual pv2p gauge

renderGauge(
  
  gauge(100 - biden_actual_pv2p,
        min = 0, 
        max = 100,
        gaugeSectors(
          success = c(50.1, 100), 
          warning = c(50, 50), 
          danger = c(0, 49.9)), 
        label = "Two-Party Popular Vote \nfor Trump")
  
)

```


Row
-------------------------------------------------

### Predicted Electoral Votes

```{r pred_ev_gauge}

# Biden predicted EV gauge

renderGauge(

  gauge(273,
        min = 0, 
        max = 538,
        gaugeSectors(
          success = c(270, 538), 
          warning = c(269, 269), 
          danger = c(0, 268)), 
        label = "Electoral Votes \nfor Biden")
)

# Trump predicted EV gauge

renderGauge(
  
  gauge(265, 
        min = 0, 
        max = 538,
        gaugeSectors(
          success = c(270, 538),
          warning = c(269, 269), 
          danger = c(0, 268)),
        label = "Electoral Votes \nfor Trump")
)
```

### Predicted Two-Party Popular Vote Share

```{r predicted_pv2p_gauge}

biden_pred_pv2p <- 52.720

# Biden predicted pv2p gauge

renderGauge(
  
  gauge(biden_pred_pv2p,
        min = 0, 
        max = 100,
        gaugeSectors(
          success = c(50.1, 100), 
          warning = c(50, 50), 
          danger = c(0, 49.9)), 
        label = "Two-Party Popular Vote \nfor Biden")
  
)

# Trump predicted pv2p gauge

renderGauge(
  
  gauge(100 - biden_pred_pv2p,
        min = 0, 
        max = 100,
        gaugeSectors(
          success = c(50.1, 100), 
          warning = c(50, 50), 
          danger = c(0, 49.9)), 
        label = "Two-Party Popular Vote \nfor Trump")
  
)


```


Maps
================================================

For the initial tab, the predicted results are displayed on the left and the actual results are displayed on the right. Hover over each state to see the state name and the exact number. Notice that the colors on the error map are almost the inverse of the actual electoral map. This means that Donald Trump performed better than predicted in the more liberal states and Joe Biden performed better than predicted in the more conservative states. I discuss this in further detail and potential hypothesis for this in my full-length [model reflection](https://kayla-manning.github.io/gov1347/posts/reflection/reflection.html) on my blog.

Maps {.tabset}
-------------------------------------------------

### Predicted vs. Actual Two-Party Vote Shares

```{r pred_map, warning=FALSE}

# creating the map with the predicted vote shares

pred_map <- ggplotly(pred_compare %>% 
  left_join(us_map, by = c("state" = "region")) %>% 
  ggplot(aes(long, lat, group = group, text = paste0("Biden's Predicted Two-Party \nVote Share in ", 
                                                     state, ": ", round(pred_pv2p, 3), "%"))) +
  geom_polygon(aes(fill = pred_pv2p)) +
  scale_fill_gradient2(low = "#BE1E26", high = "#4B5973",
                       midpoint = 50,
                       breaks = c(30, 50, 70),
                       labels = c(30, 50, 70),
                       limits = c(25, 75)) +
  coord_map() +
  labs(x = "",
       y = "",
       fill = "Biden's Two-Party \nVote Share",
       title = "Predicted Results") +
  theme_hodp() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.text.x = element_blank()),
  tooltip = "text")

# creating the map with the actual vote shares

actual_map <- ggplotly(pred_compare %>% 
  left_join(us_map, by = c("state" = "region")) %>% 
  ggplot(aes(long, lat, group = group, text = paste0("Biden's Actual Two-Party \nVote Share in ", 
                                                     state, ": ", round(actual_pv2p, 3), "%"))) +
  geom_polygon(aes(fill = actual_pv2p)) +
  scale_fill_gradient2(low = "#BE1E26", high = "#4B5973",
                       midpoint = 50,
                       breaks = c(30, 50, 70),
                       labels = c(30, 50, 70),
                       limits = c(25, 75)) +
  coord_map() +
  labs(x = "",
       y = "",
       fill = "Biden's Two-Party \nVote Share",
       title = "Predicted vs. Actual Results") +
  theme_hodp() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.text.x = element_blank()),
  tooltip = "text")

# combining the two maps into one plot

renderPlotly(
  subplot(pred_map, actual_map)
)

```

### Model Errors

```{r error_map}

# this map shows how far off my predictions were from the actual results in each
# state... it's essentially a different way of displaying the info shown in the
# other two maps but in a single map

renderPlotly(
  ggplotly(pred_compare %>% 
    left_join(us_map, by = c("state" = "region")) %>% 
    ggplot(aes(long, lat, group = group, text = paste0("Difference between actual and \npredicted vote shares in ", 
                                                       state, ": ", round(diff, 3), "%"))) +
    geom_polygon(aes(fill = diff)) +
    scale_fill_gradient2(low = my_red, mid = "white", high = my_blue,
                         midpoint = 0, limits = c(-14, 14), labels = c(-10, 0, 10),
                         breaks = c(-10, 0, 10)) +
    coord_map() +
    labs(x = "",
         y = "",
         fill = "Difference Between Biden's \nActual and Predicted \nTwo-Party Vote Share",
         title = "Forecast Error") +
    theme_hodp() +
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.text.x = element_blank()),
    tooltip = "text") %>% 
   layout(annotations = 
            list(x = 1, y = -0.1, text = "Red indicates that Trump received a greater share of votes than predicted.", 
                 showarrow = F, xref='paper', yref='paper', 
                 xanchor='right', yanchor='auto', xshift=80, yshift=40,
                 font=list(size=15))
  )
)

```


##

```{r error_table, max.height='100px'}

# creating a table that shows the actual and predicted vote shares for each
# state so that people can take a closer look at the actual numbers behind the
# maps

pred_compare %>% 
  mutate(actual_pv2p = paste0(round(actual_pv2p, 3), "%"),
         pred_pv2p = paste0(round(pred_pv2p, 3), "%"),
         diff = paste0(round(diff, 3), "%")) %>% 
  select(-2)  %>% 
  kable(col.names = c("State", "Actual Two-Party \nPopular Vote",
                "Predicted Two-Party \nPopular Vote", "Actual - Predicted")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```


Simulation Results
=================================================

Nationwide Simulations {.tabset}
-------------------------------------------------

### Simulated Electoral Counts 

```{r national_ev}

# plot that shows the electoral counts in each of my simulated elections from my
# forecast

ev_sims_plot <- ev_sims %>% 
  pivot_longer(cols = 2:3, names_to = "candidate", values_to = "ev") %>% 
  filter(candidate == "trump_ev") %>% 
  mutate(result = case_when(ev < 269 ~ "Biden",
                            ev >= 270 ~ "Trump",
                            ev == 269 ~ "Tie")) %>% 
  ggplot(aes(ev, text = paste0("Probability: ", round(after_stat(count / sum(count)), 3)))) +
    geom_density(aes(ev, ..density..), 
                 inherit.aes = FALSE) +
    geom_histogram(aes(y = after_stat(count / sum(count)), fill = result), 
                   binwidth = 1, position = "identity", alpha = 0.75) +
    labs(title = "Simulated Electoral Outcomes \nCompared to Actual Result",
         x = "Donald Trump's Electoral Votes",
         y = "Proportion of Simulations",
         fill = "Winner",
         subtitle = "Donald Trump Wins the Electoral College in 35.074% of Simulations") +
    geom_vline(xintercept = 232, size = 1)  +
  scale_fill_manual(breaks = c("Biden", "Trump", "Tie"),
                    values = c(my_blue, my_red, "gray"),
                    labels = c("Biden", "Trump", "Tie")) +
  theme_hodp()

renderPlotly(
  ggplotly(ev_sims_plot, tooltip = "text") %>% 
    add_annotations(text = "Actual Outcome", x = 232, y = 0.015, textangle = 270)
)
    

```

### Simulated Two-Party Vote Share

```{r nationwide_pv2p}

# plot that shows the two-party popular vote for each of the simulations from my
# forecast

pv2p_sims_plot <- sims %>% 
  mutate(group = rep(1:10000, times = 50)) %>% 
  group_by(group) %>% 
  summarize(total_dem = sum(sim_dvotes_2020),
         total_rep = sum(sim_rvotes_2020),
         d_pv2p = total_dem / (total_dem + total_rep),
         r_pv2p = 1 - d_pv2p,
         .groups = "drop") %>% 
  pivot_longer(cols = 4:5, names_to = "party") %>% 
  ggplot(aes(value, fill = party)) +
  geom_histogram(aes(y = after_stat(count / sum(count))), bins = 1000, position = "identity", alpha = 0.95) +
  geom_vline(xintercept = biden_actual_pv2p / 100) +
  geom_vline(xintercept = 1 - biden_actual_pv2p / 100) +
  theme_minimal() +
  scale_fill_manual(breaks = c("d_pv2p", "r_pv2p"),
                    labels = c("Biden", "Trump"),
                    values = c(my_blue, my_red)) +
  labs(title = "Simulated Popular Vote Shares \nCompared to Actual Results",
       x = "Predicted Share of the Two-Party Popular Vote",
       y = "Count",
       subtitle = "Donald Trump Wins the Popular Vote in 0.049% of Simulations",
       fill = "Candidate") +
  theme_hodp()

renderPlotly(
  ggplotly(pv2p_sims_plot) %>% 
    add_annotations(text = "Trump Actual Outcome", x = 0.48, y = 0.002, textangle = 270) %>% 
    add_annotations(text = "Biden Actual Outcome", x = 0.518, y = 0.002, textangle = 270)
)  
```

State-Level Predictions
=================================================

State-Level Predicted Vote Shares and Probabilities of Victory {.tabset}
-------------------------------------------------

### Vote Shares

```{r state_predicted_pv}

# this displays the state-level pv2p point predictions, categorized by red,
# blue, and battleground states

selectInput("state_type_pv",
            "State Category:",
            types %>% pull(type) %>% unique())
renderPlotly(
  state_voteshares(input$state_type_pv) %>%
    layout(margin = list(b = 90))
)

```

### Win Probabilities

```{r state_win_probs}

# this shows each candidate's probabilities of winning each state, calculated by
# the proportion of simulations that each candidate won that state

selectInput("state_type_win",
            "State Category:",
            types %>% pull(type) %>% unique())
renderPlotly(
  state_win_probs(input$state_type_win) %>%
    layout(margin = list(b = 90))
)
```

-------------------------------------------------

### State-Level Two-Party Vote Share Predictions

```{r state_sim_histograms}

# This show's each state's pv2ps generated in the forecast simulations

selectInput("state", 
            "State:",
            sims %>% pull(state) %>% unique() %>% sort())
renderPlotly(
  pv2p_plot(input$state) %>%
    layout(margin = list(b = 90))
)

```



State-Level Results
=================================================

### State-Level Error

```{r error_scatterplot}

# this displays the actual and the model's predicted vote share on a
# scatterplot, with each point representing a single state

error <- pred_compare %>% 
  mutate(incorrect = ifelse(state %in% c("AZ", "NV", "GA"), TRUE, FALSE)) %>% 
  ggplot(aes(pred_pv2p, actual_pv2p,
             text = paste0("Biden Overperformed in ", state, " by ", round(diff, 3), "%"),
             color = diff, shape = incorrect)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  labs(y = "Actual Two-Party Popular Vote Share",
       title = "State-Level Model Error",
       x = "Predicted Two-Party Popular Vote Share",
       color = "Error = \nActual - Predicted",
       shape = "Incorrect Prediction?") +
  theme_hodp() +
  scale_color_gradient2(low = "#BE1E26", mid = "#E2DDDB", high = "#4B5973")

renderPlotly(
  ggplotly(error, tooltip = c("text", "x", "y", "shape"))
)

```

Changes Over Time
-------------------------------------------------

### 

```{r change_linegraph}

# this linegraph shows how each candidate's pv2p changed in each state as more
# and more votes were counted on election night and beyond

selectInput("line_state", 
            "State:",
            sims %>% pull(state) %>% unique() %>% sort())

renderPlotly(
  
  ggplotly(changes %>% 
    filter(state == input$line_state) %>% 
    ggplot(aes(timestamp, pv2p, color = candidate,
               text = paste0(candidate, " had ", round(pv2p * 100, 3), 
                             "% of the vote share on ", timestamp,
                             " in ", state),
               group = candidate)) +
    geom_line() +
    geom_hline(aes(yintercept = state_pred_pv2p(input$line_state, "biden")), 
               linetype = "dashed", color = my_blue) +
    geom_hline(aes(yintercept = state_pred_pv2p(input$line_state, "trump")), 
               linetype = "dashed", color = my_red) +
    labs(y = "Two-Party \nPopular Vote Share",
         title = "Changes in Vote Share After Election Day",
         x = "",
         color = "Candidate",
         shape = "Incorrect Prediction?") +
    theme_hodp() +
    scale_color_manual(values = c(my_blue, my_red, my_blue, my_red)),
  tooltip = "text",
  height = 350)
  
)


```

